---
title: "STAT_495_Project"
author: "Noah Gallagher"
date: "4/15/2021"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in Data

```{r}
library(tidyverse)

#Reading in data
hotel.data <- read_csv('hotel_bookings.csv')
```

## Exploratory Data Analysis

First we will look at the raw data values.
```{r}
glimpse(hotel.data)

head(hotel.data)
```

Next we can compute the summary statistics
```{r}
library(skimr)
skim(hotel.data)
```

Now we can start Data Visualization.
```{r}
library(gridExtra)
# visualizing week number
a = ggplot(hotel.data, aes(x= arrival_date_week_number))+
  geom_histogram(binwidth = 8)+ 
  labs(title = "arrival week number")

# arrival date year
b = ggplot(hotel.data, aes(x= arrival_date_year))+
  geom_histogram(binwidth=)+ 
  labs(title = "arrival date year")

# arrival day of month
c = ggplot(hotel.data, aes(x= arrival_date_day_of_month))+
  geom_histogram(binwidth = 8)+ 
  labs(title = "arrival day of month")

# stay in weekend nights
d = ggplot(hotel.data, aes(x= stays_in_weekend_nights))+
  geom_histogram(binwidth = 3)+ 
  labs(title = "num stays in weekend nights")

# stay in week nights
e = ggplot(hotel.data, aes(x= stays_in_week_nights))+
  geom_histogram(binwidth = 3)+ 
  labs(title = "num stays in week nights")

# customer type
f = ggplot(hotel.data, aes(x= customer_type))+
  geom_bar()+ 
  labs(title = "customer types")

grid.arrange(a, b, c, ncol=1, nrow=3)

grid.arrange(d, e, f, ncol=1, nrow=3)
```

## Question: When is the best time of year to book a hotel room?
```{r}
# years 2015-2017
# months july 2015-august 2017
stats<- hotel.data%>%
  summarize(
    week_mean = mean(arrival_date_week_number),
    day_month_mean = mean(arrival_date_day_of_month)
  )
stats

most_freq<-hotel.data%>%
  group_by(arrival_date_month)%>%
  summarize(
    count=n()
  )%>%
  arrange(desc(count))%>%
  head(1)
most_freq

least_freq<-hotel.data%>%
  group_by(arrival_date_month)%>%
  summarize(
    count=n()
  )%>%
  arrange(count)
least_freq

```

Guests typically arrive on the 15th on the month.
Guests typically arrive the 27th week of the year.
Guests typically arrive in August.

Least busy: January
Most busy: August

## Question: How far in advance do people make bookings? Are they more inclined to cancel?
Lead time represents the number of days that elapsed between the entering date of the booking into the PMS and the arrival date.

```{r}
attach(hotel.data)

# visualizing lead time FOR ALL
ggplot(hotel.data, aes(x= lead_time))+
  geom_histogram(bins=5)+
  labs(title = "Booking Lead Time", x = "Lead Time (in days)", y = "Number of Bookings")

# visualizing cancellation rate
ggplot(hotel.data, aes(x= is_canceled))+
  geom_bar(bins=5)+
  labs(title = "Cancellations", subtitle = '0 = NO, 1 = YES')

#average lead time
cat("the average lead time is: ", mean(lead_time))

# Is the higher lead time associated with higher cancellation?
cat("\nthe correlation btw lead time and cancellation is: ",cor(lead_time, is_canceled))


cancellation<-hotel.data%>%
  select(lead_time, is_canceled)
pairs(cancellation)

mean(is_canceled)
```
Guests typically book hotels 104 days prior to their arrival.

Lead time and Cancellations have a positive correlation of 0.29


## Seperating the data into City hotels and Resport hotels
```{r}
seperate_hotels <- hotel.data%>%
  group_by(hotel)%>%
  summarise(
    entries = n(),
    #week_mean = mean(arrival_date_week_number),
    #day_month_mean = mean(arrival_date_day_of_month),
    avg_lead = mean(lead_time),
    #corr = cor(lead_time, is_canceled),
    avg_can = mean(is_canceled)
  )%>%
  mutate(
    percent_cancell = round(avg_can*100, digits = 2)
  )
seperate_hotels
```
Cancellation Rate:
  - City Hotels have 41.73% cancellation rate.
  - Resort Hotels have 27.76% cancellation rate.

Correlation:
  - City Hotels - cancellation and lead time is 31% correlated (positive)
  - Resort Hotels  - cancellation and lead time is 23% correlated (positive)


## Bootstrapping

```{r}
library(infer)
set.seed(99999)

#Size 1,000 City
city_sample <- hotel.data%>%
  filter(hotel == 'City Hotel')%>%
  rep_sample_n(size = 1000, reps = 50)
city_sample

#Size 1,000 Resort
resort_sample <- hotel.data%>%
  filter(hotel == 'Resort Hotel')%>%
  rep_sample_n(size = 1000, reps = 50)
resort_sample

# getting stats from city sample
city_sample_stats <- city_sample%>%
  group_by(replicate)%>%
  summarise(
    week_mean = mean(arrival_date_week_number),
    day_month_mean = mean(arrival_date_day_of_month),
    avg_lead = mean(lead_time),
    corr = cor(lead_time, is_canceled),
    avg_can = mean(is_canceled)
  )%>%
  mutate(
    percent_cancell = round(avg_can*100, digits = 2)
  )
city_sample_stats

mean(city_sample_stats$avg_lead)
mean(city_sample_stats$corr)
mean(city_sample_stats$percent_cancell)

# getting stats from resort sample
resort_sample_stats <- resort_sample%>%
  group_by(replicate)%>%
  summarise(
    week_mean = mean(arrival_date_week_number),
    day_month_mean = mean(arrival_date_day_of_month),
    avg_lead = mean(lead_time),
    corr = cor(lead_time, is_canceled),
    avg_can = mean(is_canceled)
  )%>%
  mutate(
    percent_cancell = round(avg_can*100, digits = 2)
  )
resort_sample_stats

mean(resort_sample_stats$avg_lead)
mean(resort_sample_stats$corr)
mean(resort_sample_stats$percent_cancell)

# plotting avg lead time for city sample
ggplot(city_sample_stats, aes(x= avg_lead))+
  geom_histogram(bins=10)+
  labs(title = "Lead time Booking for Bootstrapped City Hotels", x = "Average Lead Time")

# plotting avg lead time for resort sample
ggplot(resort_sample_stats, aes(x= avg_lead))+
  geom_histogram(bins=10)+
  labs(title = "Lead time Booking for Bootstrapped Resort Hotels", x = "Average Lead Time")
```

## Prediction from Samples

Q: Can we predict the arrival date for 2 adults with 2 children.
```{r}
library(moderndive)

# Prediction using City Hotel data
city_predict <- lm(arrival_date_day_of_month ~ adults + children, data=city_sample)
get_regression_table(city_predict)

# Prediction using Resort Hotel data
resort_predict <- lm(arrival_date_day_of_month ~ adults + children, data=resort_sample)
get_regression_table(resort_predict)

#Prediction using ALL hotels - this is on teh presentation
hotel_arrival <- lm(arrival_date_day_of_month ~ adults + children + babies, data=hotel.data)
get_regression_table(hotel_arrival)

# Prediction using Resort Hotel data
resort_predict_ad2_ch2_resort <- lm(is_canceled~lead_time+previous_bookings_not_canceled, data=resort_sample)
get_regression_table(resort_predict_ad2_ch2_resort)

# number of special requests
attach(hotel.data)
spec_req <- lm(total_of_special_requests ~ adults + children + babies)
get_regression_table(spec_req)

# number of cancellation
hotel_cancelled<- lm(is_canceled ~ lead_time + previous_bookings_not_canceled, data=hotel.data)
get_regression_table(hotel_cancelled)
```
City Hotel: Prediction date of 15.917 + 2(-0.062) + 2(0.354) = 16.5
- Arrival will be on the 17th

Resort Hotel: Prediction date of 15.756 + 2(0.018) + 2(0.387) = 16.566 = 17
- Arrival will be on the 17th


## Regression : lead time vs. arrival week number

Using regression to form a line that tells us the relationship between how far in advance guests book their stays and their arrival week number.
```{r}
ggplot(hotel.data, aes(y = arrival_date_week_number, x = lead_time))+
  geom_point(shape = 1) +
  geom_smooth(method = lm) +
  #geom_jitter(shape = 1)+
  labs(y='arrival week #',x='lead time')

```
Findings: People that arrive later in the year, often book further in advance.

